Solar wind, F10.7, and geomagnetic activity relationship to the equatorial plasma mass density at geosynchronous orbit

### 2016-09-02 paper/paper.tex
Updated paper figures, added denton notes

\begin{abstract}
We consider two types of events, identified by decreases in $D_{st}$ below a threshold value and increases in the the equatorial mass density at geosynchronous altitudes, $\rho_{eq}$, above a threshold value using the \cite{Takahashi2010} dataset.  From the $D_{st}$ events and 1-day averages, we find that there is a statistically weak and small-amplitude difference between $\rho_{eq}$ on the day of the event and the days before and after.  When hourly averages are considered, a signficant peak is found to occur six hours after event onset, and the primary factor that determines the post-onset peak amplitude in $\rho_{eq}$ is elevated $F10.7$.  In addition, for hourly averages, $\rho_{eq}$ following the onset of a $D_{st}$ event does not depend on the north-south component of the interplanetary magnetic field, $B_z$ near the time of onset.  From the $\rho_{eq}$ events, we find a weak dependence on $B_{z}$ after the onset of an event, with higher average $B_{z}$ four hours after the event onset corresponding to larger $\rho_{eq}$ values 24-36 hours after onset.  In contrast, the average of $B_z$ four hours prior to $\rho_{eq}$ events does not have a statistically significant impact on $\rho_{eq}$ after onset.
\end{abstract}
%\vspace{2em}
%\hrule
%\vspace{2em}
% \end{@twocolumnfalse}
%  ]

@@

\cite{Takahashi2006} estimated magnetospheric mass density using measurements from the CRRES satellite during a 73-day period in 1991 when CRRES was in an elliptical, near equatorial orbit and primarily beyond the nominal plasmapause. They found that the local average ion mass, $M$, had some correlation with geomagnetic activity, with more negative hourly $D_{st}$ values corresponding to higher $M$ and 1.5- and 3-day averages of $K_p$ corresponding to higher $M$ (3-hour averages of $K_p$ had little visual correlation with $M$). The mass density estimates were obtained from CRRES observations in the MLT range of [12:00, 18:00] with most observations at $L$ between 5 and 7 $R_E$.

\cite{Denton2006} found a weak relationship between $D_{st}$ and and $K_p$ and the field line distribution of plasma mass density, $\rho$, using observations of the first three toroidal Alfv\'en harmonic frequencies for $L$ in the range of 6-8~$R_E$.  A more pronounced peak in the distribution along a field line near the equator was found for lower $D_{st}$ and larger $K_p$, but the equatorial value of $\rho$ was found to be similar for $D_{st}$ bins of [-142, -31] and [-31, 37] nT and $K_p$ bins of [1.5,3.4] and [3.4, 5.9].

@@

\cite{Takahashi2010} noted that downward drops in the $D_{st}$ index coincided with significant changes in $\rho_{eq}$ for $L$ near 6.8~$R_E$. For five storms, two had $\rho_{eq}$ spikes after $D_{st}$ minimum, two had $\rho_{eq}$ spikes before the drop, and one showed little change in $\rho_{eq}$.  A key result was that when daily-averaged measurements were considered, the epoch average of $D_{st}$ exhibited for storms with a minimum $D_{st} < -50$~nT showed an enhancement in $\rho_{eq}$ the same day as minimum $D_{st}$ in the epoch averages.

@@

The above results indicate that (1) the mass density in the at geosynchronous distances is best correlated with $F10.7$ and (2) there exists a much weaker statistical relationship between mass density and the geomagnetic activity indices $K_p$ and $D_{st}$.  In this work we consider the dependence of mass density estimates in the \cite{Takahashi2010} data set and attempt to identify and statistically characterize the relationship between geomagnetic activity with mass density at geosynchronous distances.  In addition, we attempt to identify solar wind variables that may drive the changes.

@@

The parameters $\rho_{eq}$ and $F_{10.7}$ are from the dataset of \cite{Takahashi2010}; data are available from 1980 through 1991 from GOES 2, 3, 5, 6 and 7.  All other parameters used are from \cite{Kondrashov2014ReconstructionOfGaps}, which has coverage from 1972 through 2013 and are on a 1--hour time grid. $\rho_{eq}$ is the inferred equatorial mass density based on the 3rd harmonic torodial frequency of magnetic field measurements as described in \cite{Takahashi2010}.  The smallest cadence for $\rho_{eq}$ values is 10 minutes.  To compute an hourly median over the same interval for which the solar wind parameters were averaged, the median of all $\rho_{eq}$ values in a given hour window was used. GOES 6 had the longest span of available (May 1983 to August 1991) data and the primary results presented are for GOES 6 while measurements from other satellites were used for consistency checks.

@@

Two types of events are considered. The first is a drop in the $D_{st}$ index below the threshold value of -50~nT. These types of events are considered in order to determine if there is a well-defined $\rho_{eq}$ dependence on geomagnetic activity/storms as indicated by the $D_{st}$ index.  It is expected that the plasmasphere will exhibit two responses near these threshold crossings.  First is the movement of the plasmapause earthward (\cite{LemaireEarthsPlasmasphere}) and the second is a possible change in density due to magnetospheric processes.

The second type of event is an increase in $\rho_{eq}$ above a threshold of 20~amu/cm$^3$.  It is expected that some of these increases will typically occur after period of quiet geomagnetic activity.  In this case, the plasmapause boundary may move outward past geosynchronous orbit and the increase will be due to measurements being made inside the plasmapause.  A second possible reason for the increase could be due to the same mechanism associated with increases that occur during enhanced geomagnetic activity.

@@

Our first analysis uses GOES-6 data only from the time interval 1989-1991, which corresponds with the interval used in the epoch analysis of \cite{Takahashi2010}. In this interval, there were 75 $D_{st}$ events where $D_{st}$ remained below the $-50$~nT threshold for at least 12 hours. Only events with an onset time between 06:00 and 12:00 MLT were considered. The first hour of a local minimum in $D_{st}$ following a threshold crossing defines the zero epoch time, and for each event, $4\cdot24$ hours were considered before and $8\cdot24$ hours after onset. For multiple detected threshold crossings within 24 hours of each other, the first crossing was selected as the event onset because visual inspection showed the subsequent crossings to typically be from brief downward deviations during the recovery period of a geomagnetic storm. To compute the epoch averages on a daily time scale, the median value of all available measurements for all events centered on a window of $\pm 12$ hours of the epoch zero hour was computed, and these averaging windows were shifted in increments of $24$~hours. Similar results are obtained if we first reduce each event time series to have a 1-day cadence by computing medians for each event in 1-day bins and then compute the medians across the averaged event time series on each epoch day.

The stack plot in the upper panel of Figure \ref{fig:DailyAveragedDstEvents} shows the epoch averages for the 75 events from 1989-1991.  The minimum $D_{st}$ median is $-75$~nT.  Consistent with \cite{Takahashi2010}, $D_{st}$ events correspond to elevated $\rho_{eq}$ before and after the event, although the magnitude of increase observed here is $4.5$ amu/cm$^3$ instead of the increase of ${\sim} 10$ amu/cm$^3$ found in Figure 11 of \cite{Takahashi2010}.  The vertical green bars in the $\rho_{eq}$ plot show the number of $\rho_{eq}$ values that were used to compute the medians and red error lines.  Error lines for each parameter are the standard deviation of the values used in computing the median divided by the square root of the number of values. For the non-$\rho_{eq}$\ parameters, the number of measurements used in computing the medians is typically equal to the number of events, which is much larger than the number of $\rho_{eq}$\ values as all available measurements were used instead of restricting to only values where a $\rho_{eq}$ value also existed.

In the lower panel of Figure \ref{fig:DailyAveragedDstEvents}, it is shown that the trends seen in the analysis of 1989-1991 also hold for the entire span of measurements from GOES 6 (1983-1991) - a small elevation on the day of the event relative to the day before and day after, with median averages after the day of the event being slightly lower than that before.  This trend is also seen for GOES 7 (1987-1991), is much less significant for GOES 5 (1983-1987), and much more significant for GOES 2 (1981-1983).  We note here that this trend has a relationship with the average value of F10.7 during each respective coverage period, with the most significance found during the 1981-1983 time interval when F10.7 was consistently elevated.

@@

A similar hypothesis test run for the bottom panel of Figure~\ref{fig:DailyAveragedDstEvents} yields a maximum confidence level of $\sim$85\%.

These tests indicate that there is not a statistically significant difference in medians of daily-averaged $\rho_{eq}$ on the day before and day after a threshold crossing in $D_{st}$.  In addition, the observed differences are on the order of only 20\% of the value of $\rho_{eq}$ on the onset day, which is similar to the variation in the observations over the displayed epoch time interval.

@@

We have also considered the hourly averages from the events shown in the bottom panel of Figure~\ref{fig:DailyAveragedDstEvents}, which shown in the top panel of Figure~\ref{fig:HourlyAveragedDstEvents}.  In the time period of elevated mass density within 6 hours of the start of the event, there are only $\sim$4 samples per hourly interval and the peak value of $\sim$70 amu/cm$^3$ is associated with a single observation.

To increase the number of events, we have removed the constraint that the events must occur in the MLT range of 06:00-12:00 and the constraint that $D_{st}$ remained below the threshold value for 12 hours or longer.  The result is shown in the bottom panel of Figure~\ref{fig:HourlyAveragedDstEvents}. The two variations from a baseline (one downwards before onset and one upwards after onset) were tested for significant deviation from two baselines, one 24-12 hours before onset, one 16-48 hours after onset. The probability of the downward variation coming from the same distribution as the data from the left baseline is 1.06\%, whereas the probability of the right upward variation coming from the same distribution of the data for the right baseline is 0.17\%.  This significant result is also maintained for GOES 2, 5, and 7.  This indicates that there are statistically significant local variation in $\rho_{eq}$ within 6 hours of onset when hourly averages are considered.  The signficance of the upward variation result is consistent with previous observations of large mass density increases in the plasma trough region during large geomagnetic storms (\cite{Yao2008}; \cite{Takahashi2010}).

@@

To determine if there is an IMF $B_{z}$ dependence, the epoch curves in the bottom panel of Figure~\ref{fig:HourlyAveragedDstEvents} for $\rho_{eq}$ and $B_{z}$ were separated into two parts depending whether $B_z$ was above or below its average four hours before and including the onset hour and four hours after and including the hour of onset; the result is shown in Figure~\ref{fig:BzBinnedDstEvents}. Any statistically significant difference at a 95\% confidence level between the two parts is indicated by a green dot on the horizontal axis using a Wilcoxon rank-sum test. For verification, the significance values were also tested with the same bootstrap method described previously and found to be similar.  (At this confidence level, false positives are expected to randomly occur 5\% of the time, or 3.6 times for 72 values.)  These two plots indicate that there is not a statistically significant dependence on IMF $B_{z}$ around the time of the onset of a $D_{st}$ event.

The process for creating Figure~\ref{fig:BzBinnedDstEvents} was repeated except using the sort variable $F10.7$ instead of $B_z$, and the result is shown in Figure~\ref{fig:F107BinnedDstEvents}. The difference between the epoch curves is statistically significant at all times, as expected due to the high correlation between $\rho_{eq}$ and $F10.7$, but elevated $F10.7$ is associated with a peak in the $\rho_{eq}$ curve approximately 5 hours after onset. The values contributing to the peak were compared to a baseline of all values 16-48 hours after onset via a Wilcoxon rank-sum test and were found to have significantly different medians at the 99\% confidence level.  This indicates the possibility that there must be a high pre-existing baseline level of $\rho_{eq}$ in order for an increase in $\rho_{eq}$ to be observed for a $D_{st}$ event.  This result is also consistent with the the analysis of Figure~3 when repeated using measurements from GOES satellites taken during different parts of the solar cycle; the significance of the differences between the days around onset were largest when the measurements were take during time intervals when F10.7 was consistently elevated.

@@

Figure~\ref{fig:BzBinned} shows the the events of Figure~\ref{fig:HourlyAveragedRhoEvents} when separated by $B_z$ in the same manner as Figure~\ref{fig:BzBinnedDstEvents}.  The largest difference occurs when the separation is performed after onset, with more positive $B_z$ leading to a larger peak in $\rho_{eq}$ than more negative $B_{z}$.  The rate of change of $\rho_{eq}$ also appears to be insensitive to $B_z$ for in the time windows used to determine the separation.

@@

Figure~\ref{fig:F107Binned} shows how the epoch average curve of $\rho_{eq}$ in Figure~\ref{fig:HourlyAveragedRhoEvents} depends on $F10.7$.  Statistically significant differences occur independent of if the separating factor the average of $F10.7$ before or after onset, as expected because the cadence of $F10.7$ is one day and its time scale of variation is longer than that of $\rho_{eq}$.  Events with high $F10.7$ are associated with higher $\rho_{eq}$ before and after event onset but similar peak values.  As was the case for $B_z$, there is an approximately 12--hour time span of linear growth of $\rho_{eq}$ near the time of onset but because the initial starting $\rho_{eq}$ is higher for high $F10.7$, the associated growth rate is lower.

@@

On average, around the time of large geomagnetic events, there is a statistically weak enhancement of daily-averaged $\rho_{eq}$ measurements. For hourly averages, $\rho_{eq}$ shows enhancements over a 6-hour post-onset window that contains too few events for a statistical analysis.  When the constraint of extended geomagnetic events (greater than 12 hours below a threshold value) and the MLT of the observation of $\rho_{eq}$ (06:00-12:00) is removed to allow for more events, statistically significant variations are found, with a peak occurring approximately six hours after the event onset.  For the hourly-averaged measurements, the average of $B_{z}$ around onset time did not have a significant impact on $\rho_{eq}$ after onset.

It was shown that a key factor in determining if large $\rho_{eq}$ enhancements will be observed is elevated $F10.7$.  When geomagnetic events were separated according to if they occurred when $F10.7$ was low or high, significant enhancements in $\rho_{eq}$ above a baseline value were only observed under high $F10.7$ conditions.

For $\rho_{eq}$, more positive $B_{z}$ after onset corresponded to higher values of subsequent $\rho_{eq}$, which indicates that a substantial portion of the events may due to the apparent mass refilling mechanism described by \cite{Denton2016}.  In order to separate this mechanism from an enhancement due to magnetospheric activity, additional information about the location of the plasmapause may be required.

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.40]{figures/F107MD27d-GOES6.eps}
\includegraphics[scale=0.40]{figures/ccplot-GOES6.eps}
\caption{Top: 27-day non-overlapping medians of $F_{10.7}$ and $\log_{10}(\rho_{eq})$ from GOES 6. Bottom: Correlation between $\log_{10}(\rho_{eq})$ and $F_{10.7}$ using medians in non-overlapping hour, day, and 27-day windows.}
\label{fig:ccplot}
\end{figure}
\clearpage

>>>>

\begin{abstract}
We consider two types of events, identified by decreases in $D_{st}$ below a threshold value and increases in the the equatorial mass density at geosynchronous altitudes, $\rho_{eq}$, above a threshold value using the \cite{Takahashi2010} dataset.  From the $D_{st}$ events and 1-day averages, we find that there is a statistically weak and small-amplitude difference between $\rho_{eq}$ on the day of the event and the days before and after.  When hourly averages are considered, a significant peak is found to occur six hours after event onset, and the primary factor that determines the post-onset peak amplitude in $\rho_{eq}$ is elevated $F10.7$.  In addition, for hourly averages, $\rho_{eq}$ following the onset of a $D_{st}$ event depends on the north-south component of the interplanetary magnetic field, $B_z$ after the time of onset, with higher average $B_{z}$ four hours after the event onset corresponding to larger $\rho_{eq}$ value 7-11 hours after onset.  From the $\rho_{eq}$ events, we find a weak dependence on $B_{z}$ after the onset of an event, with higher average $B_{z}$ four hours after the event onset corresponding to larger $\rho_{eq}$ values 24-36 hours after onset.
\end{abstract}
%\vspace{2em}
%\hrule
%\vspace{2em}
% \end{@twocolumnfalse}
%  ]

@@

\cite{Takahashi2006} estimated magnetospheric mass density using measurements from the CRRES satellite during a 73-day period in 1991 when CRRES was in an elliptical, near equatorial orbit and primarily beyond the nominal plasmapause. They found that the local average ion mass, $M$, had some correlation with geomagnetic activity, with more negative hourly $D_{st}$ values corresponding to higher $M$ and larger 1.5- and 3-day averages of $K_p$ corresponding to higher $M$ (3-hour averages of $K_p$ had little visual correlation with $M$). The mass density estimates were obtained from CRRES observations in the MLT range of [12:00, 18:00] with most observations at $L$ between 5 and 7 $R_E$.

\cite{Denton2006} found a weak relationship between $D_{st}$ and $K_p$ and the field line distribution of plasma mass density, $\rho$, using observations of the first three toroidal Alfv\'en harmonic frequencies for $L$ in the range of 6-8~$R_E$.  A more pronounced peak in the distribution along a field line near the equator was found for lower $D_{st}$ and larger $K_p$, but the equatorial value of $\rho$ was found to be similar for $D_{st}$ bins of [-142, -31] and [-31, 37] nT and $K_p$ bins of [1.5,3.4] and [3.4, 5.9].

@@

\cite{Takahashi2010} noted that downward drops in the $D_{st}$ index coincided with significant changes in $\rho_{eq}$ for $L$ near 6.8~$R_E$. For five storms, two had $\rho_{eq}$ increases after $D_{st}$ minimum, two had $\rho_{eq}$ spikes before the minimum, and one showed little change in $\rho_{eq}$.  A key result was that when daily-averaged measurements were considered, the epoch average of $D_{st}$ for storms with a minimum $D_{st} < -50$~nT showed an enhancement in $\rho_{eq}$ the same day as minimum $D_{st}$ in the epoch averages.

@@

The above results indicate that (1) the mass density at geosynchronous distances is best correlated with $F10.7$ and (2) there exists a much weaker statistical relationship between mass density and the geomagnetic activity indices $K_p$ and $D_{st}$.  In this work we consider the dependence of mass density estimates in the \cite{Takahashi2010} data set and attempt to identify and statistically characterize the relationship between geomagnetic activity with mass density at geosynchronous distances.  In addition, we attempt to identify solar wind variables that may drive or be associated with the changes.

@@

The parameters $\rho_{eq}$ and $F_{10.7}$ are from the dataset of \cite{Takahashi2010} data are available from 1980 through 1991 from GOES 2, 3, 5, 6 and 7.  All other parameters used are from \cite{Kondrashov2014ReconstructionOfGaps}, which has coverage from 1972 through 2013 and are on a 1--hour time grid. $\rho_{eq}$ is the inferred equatorial mass density based on the 3rd harmonic torodial frequency of magnetic field measurements as described in \cite{Takahashi2010}.  The smallest cadence for $\rho_{eq}$ values is 10 minutes.  To compute an hourly median over the same interval for which the solar wind parameters were averaged, the median of all $\rho_{eq}$ values in a given hour window was used. GOES 6 had the longest span of available (May 1983 to August 1991) data and the primary results presented are for GOES 6 while measurements from other satellites were used for consistency checks.

@@

Two types of events are considered. The first is a drop in the $D_{st}$ index below the threshold value of -50~nT. These types of events are considered in order to determine if there is a well-defined $\rho_{eq}$ dependence on geomagnetic activity as indicated by the $D_{st}$ index.  It is expected that the plasmasphere will exhibit two responses near the time of these threshold crossings.  First is the movement of the plasmapause earthward (\cite{LemaireEarthsPlasmasphere}) and the second is a possible change in density due to magnetospheric processes.

The second type of event is an increase in $\rho_{eq}$ above a threshold of 20~amu/cm$^3$.  It is expected that some of these increases will typically occur after a period of quiet geomagnetic activity.  In this case, the plasmapause boundary may move outward past geosynchronous orbit and the increase will be due to measurements being made inside the plasmapause.  A second possible reason for the increase could be due to the same mechanism associated with increases that occur during enhanced geomagnetic activity.

@@

Our first analysis uses GOES-6 data only from the time interval 1989-1991, which corresponds to the interval used in the epoch analysis of \cite{Takahashi2010}. In this interval, there were 75 $D_{st}$ events where $D_{st}$ remained below the $-50$~nT threshold for at least 12 hours. Only events with an onset time between 06:00 and 12:00 MLT were considered. The first hour of a local minimum in $D_{st}$ following a threshold crossing defines the zero epoch time, and for each event, $4\cdot24$ hours were considered before and $8\cdot24$ hours after onset. For multiple detected threshold crossings within 24 hours of each other, the first crossing was selected as the event onset because visual inspection showed the subsequent crossings to typically be from brief downward deviations during the recovery period of a geomagnetic storm. To compute the epoch averages on a daily time scale, the median value of all available measurements for all events centered on a window of $\pm 12$ hours of the epoch zero hour was computed, and these averaging windows were shifted in increments of $24$~hours. Similar results are obtained if we first reduce each event time series to have a 1-day cadence by computing medians for each event in 1-day bins and then compute the medians across the averaged event time series on each epoch day.

The stack plot in the upper panel of Figure \ref{fig:DailyAveragedDstEvents} shows the epoch averages for the 75 events from 1989-1991 using data from GOES 6.  The minimum $D_{st}$ median is $-75$~nT.  Consistent with \cite{Takahashi2010}, $D_{st}$ events correspond to elevated $\rho_{eq}$ before and after the event, although the magnitude of increase observed here is $4.5$ amu/cm$^3$ instead of the increase of ${\sim} 10$ amu/cm$^3$ found in Figure 11 of \cite{Takahashi2010}.  The vertical green bars in the $\rho_{eq}$ plot show the number of $\rho_{eq}$ values that were used to compute the medians and red error lines.  Error lines for each parameter have a width that is twice the standard deviation of the values used in computing the median divided by the square root of the number of values. For the non-$\rho_{eq}$\ parameters, the number of measurements used in computing the medians is typically equal to the number of events, which is much larger than the number of $\rho_{eq}$\ values as all available measurements were used instead of restricting to only values where a $\rho_{eq}$ value also existed.

In the lower panel of Figure \ref{fig:DailyAveragedDstEvents}, it is shown that the trends seen in the analysis of 1989-1991 also hold for the entire span of measurements from GOES 6 (with coverage years of 1983-1991) - a small elevation on the day of the event relative to the day before and day after, with median averages after the day of the event being slightly lower than that before.  A similar result is obtained for GOES 7 (1987-1991); the result is much less significant for GOES 5 (1983-1987) and is much more significant for GOES 2 (1981-1983).  This trend has a relationship with the average value of F10.7 during each respective coverage period, with the most significance found during the 1981-1983 time interval when F10.7 was consistently elevated.

@@

A similar hypothesis test run for the bottom panel of Figure~\ref{fig:DailyAveragedDstEvents} yields a maximum confidence level of $\sim$85\%.  For verification, the significance values were also tested using a Wilcoxon rank-sum test and were found to be similar.

These tests indicate that there is not a statistically significant difference in medians of daily-averaged $\rho_{eq}$ on the day before and day after a threshold crossing in $D_{st}$.  In addition, the observed differences are on the order of only 20\% of the value of $\rho_{eq}$ on the onset day, which is similar to the magnitude of variation in the observations over the displayed epoch time interval.

@@

We have also considered hourly averages of data for the events shown in the bottom panel of Figure~\ref{fig:DailyAveragedDstEvents}, which are shown in the top panel of Figure~\ref{fig:HourlyAveragedDstEvents}.  In the time period of elevated mass density within 6 hours of the start of the event, there are only $\sim$4 samples per hourly interval and the peak value of $\sim$70 amu/cm$^3$ is associated with a single observation.

To increase the number of events, we have removed the constraint that the events must occur in the MLT range of 06:00-12:00 and the constraint that $D_{st}$ remained below the threshold value for 12 hours or longer.  The result is shown in the bottom panel of Figure~\ref{fig:HourlyAveragedDstEvents}. The two variations from a baseline (one downwards before onset and one upwards after onset) were tested for significant deviation from two baselines, one 24-12 hours before onset, one 16-48 hours after onset. The probability of the downward variation coming from the same distribution as the data from the left baseline is 1.06\%, whereas the probability of the right upward variation coming from the same distribution of the data for the right baseline is 0.17\%.  This significant result is also maintained for GOES 2, 5, and 7.  This indicates that there are statistically significant variations in $\rho_{eq}$ within 6 hours of onset when hourly averages are considered.  The signficance of the upward variation result is consistent with previous observations of large mass density increases in the plasma trough region during large geomagnetic storms (\cite{Yao2008}; \cite{Takahashi2010}).

@@

To determine if there is an IMF $B_{z}$ dependence, the epoch curves in the bottom panel of Figure~\ref{fig:HourlyAveragedDstEvents} for $\rho_{eq}$ and $B_{z}$ were separated into two parts depending whether $B_z$ was above or below its average four hours before and including the onset hour and four hours after and including the hour of onset; the result is shown in Figure~\ref{fig:BzBinnedDstEvents}. Any statistically significant difference at a 95\% confidence level between the two parts is indicated by a green dot on the horizontal axis using a Wilcoxon rank-sum test. For the case where the separation is made according to the value of $B_{z}$ after onset, there is a visible and significant separation between the curves from 7-11 hours after onset, with more negative $B_z$ values corresponding to lower values of $\rho_{eq}$.

% (At this confidence level, false positives are expected to randomly occur 5\% of the time, or 3.6 times for 72 values.)

The process for creating Figure~\ref{fig:BzBinnedDstEvents} was repeated except using the sort variable $F10.7$ instead of $B_z$, and the result is shown in Figure~\ref{fig:F107BinnedDstEvents}. The difference between the epoch curves is statistically significant at all times, as expected due to the high correlation between $\rho_{eq}$ and $F10.7$, but elevated $F10.7$ is associated with a peak in the $\rho_{eq}$ curve approximately 5 hours after onset. The values contributing to the peak were compared to a baseline of all values 16-48 hours after onset via a Wilcoxon rank-sum test and were found to have significantly different medians at the 99\% confidence level.  This indicates the possibility that there must be a high pre-existing baseline level of $\rho_{eq}$ in order for an increase in $\rho_{eq}$ to be observed for a $D_{st}$ event.  This result is also consistent with the the analysis of Figure~3 when repeated using measurements from GOES satellites taken during different parts of the solar cycle; the significance of the differences between the days around onset were largest when the measurements were taken during time intervals when F10.7 was consistently elevated.

@@

Figure~\ref{fig:BzBinned} shows the the events of Figure~\ref{fig:HourlyAveragedRhoEvents} when separated by $B_z$ in the same manner as Figure~\ref{fig:BzBinnedDstEvents}.  The largest difference occurs when the separation is performed after onset, with more positive $B_z$ leading to a larger peak in $\rho_{eq}$ than more negative $B_{z}$.  The rate of change of $\rho_{eq}$ also appears to be insensitive to $B_z$ for the time windows used to determine the separation.

@@

Figure~\ref{fig:F107Binned} shows how the epoch average curve of $\rho_{eq}$ in Figure~\ref{fig:HourlyAveragedRhoEvents} depends on $F10.7$.  Statistically significant differences occur independent of whether the separating factor is the average of $F10.7$ before or after onset, as expected because the cadence of $F10.7$ is one day and its time scale of variation is longer than that of $\rho_{eq}$.  Events with high $F10.7$ are associated with higher $\rho_{eq}$ before and after event onset but similar peak values.  As was the case for $B_z$, there is an approximately 12--hour time span of linear growth of $\rho_{eq}$ near the time of onset but because the initial starting $\rho_{eq}$ is higher for high $F10.7$, the associated growth rate is lower.

@@

On average, around the time of large geomagnetic events, there is a statistically weak enhancement of daily-averaged $\rho_{eq}$ measurements. For hourly averages, $\rho_{eq}$ shows enhancements over a 6-hour post-onset window that contains too few events for a statistical analysis.  When the constraint of extended geomagnetic events (greater than 12 hours below a threshold value) and the MLT of the observation of $\rho_{eq}$ (06:00-12:00) is removed to allow for more events, statistically significant variations are found, with a peak occurring approximately six hours after the event onset.  For the hourly-averaged measurements, $\rho_{eq}$ averages within 7-11 hours after the onset time was lower when the average of $B_{z}$ four hours after the onset time was more positive.  This result highlights the complexity of the response of $\rho_{eq}$ to solar wind driving.  Negative values of $B_{z}$ are generally needed for a large geomagnetic event to occur and we have shown an association between large geomagnetic events and increases in $\rho_{eq}$ within six hours of the event onset, but after the onset of the event, more positive values of $B_z$ are associated with larger $\rho_{eq}$ values.

It was shown that a key factor in determining if large $\rho_{eq}$ enhancements will be observed is elevated $F10.7$.  When geomagnetic events were separated according to if they occurred when $F10.7$ was low or high, larger enhancements in $\rho_{eq}$ above a baseline value were only observed under high $F10.7$ conditions.

For $\rho_{eq}$ events, more positive $B_{z}$ after onset corresponded to higher values of subsequent $\rho_{eq}$, which indicates that a substantial portion of the events may be due to the apparent mass refilling mechanism described by \cite{Denton2016}.  In order to separate this mechanism from an enhancement due to magnetospheric activity using this type of analysis, additional information about the location of the plasmapause may be required.

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.40]{figures/F107MD27d-GOES6.eps}
\includegraphics[scale=0.40]{figures/ccplot-GOES6.eps}
\caption{Top: 27-day non-overlapping medians of $F_{10.7}$ and $\log_{10}(\rho_{eq})$ from GOES 6. Bottom: Scatter plot of $\log_{10}(\rho_{eq})$ and $F_{10.7}$ using medians in non-overlapping hour, day, and 27-day windows.}
\label{fig:ccplot}
\end{figure}
\clearpage
